

\textbf{Distributed Real-Time Systems}

Chapter 1 The Real-Time Environment

Chapter 2 Distributed Real-Time Systems

Chapter 3 Global Time

Chapter 4 Modeling Real-Time Systems

Chapter 5 Real-Time Entities and Images

Chapter 6 Fault Tolerance

Chapter 7 Real-Time Communication

Chapter 8 Time-Triggered Protocols

Chapter 9 Input and Output

Chapter 10 Real-Time Operating Systems: OSEK and AUTOSAR

Chapter 11 Real-Time Scheduling

Chapter 12 Validation

Lecturer: Prof. Dr. Jörg Friedrich, joerg.friedrich@hs-esslingen.de,
Room F 01.316

\section{Literature for this lecture}\label{literature-for-this-lecture}

\begin{longtable}[c]{@{}l@{}}
\toprule
\textbf{Text books for this lecture}\tabularnewline
{[}1.1{]}\tabularnewline
{[}1.2{]}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[c]{@{}l@{}}
\toprule
\textbf{Complementing texts}

\emph{Some books complementing the material treated in this
lecture}\tabularnewline
{[}2.1{]}\tabularnewline
{[}2.2{]}\tabularnewline
{[}2.3{]}\tabularnewline
{[}2.4{]}\tabularnewline
{[}2.5{]}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[c]{@{}l@{}}
\toprule
\textbf{Journal Articles and Web Documents}

\emph{Original journal articles and documents from the web pertaining to
this lecture}\tabularnewline
{[}3.0{]}\tabularnewline
{[}3.1{]}\tabularnewline
\tabularnewline
\bottomrule
\end{longtable}


\textbf{Chapter 1: The Real-Time Environment}
\begin{frame}
\begin{itemize}
\item  
  Definition of a real-time system.
\item  
  Simple model with operator, computer system, and controlled object.
\item  
  Introduction of distributed real-time systems.
\item  
  Hard real-time systems and soft real-time systems.
\item  
  Functional, temporal, and dependability requirements.
\item  
  Sphere of control
\item  
  Event-triggered versus time-triggered systems.
\end{itemize}
\end{frame}
\textbf{Chapter 2: Distributed Real-Time Systems}

\begin{itemize}
\item
  Distributed system architecture overview, clusters, nodes,
  communication network
\item
  Structure of node with host computer, communication network interface,
  communication controller
\item
  Event and state messages, gateways.
\item
  Concept of composability.
\item
  Event- and time-triggered communication systems.
\item
  Scalability, dependability, issues of physical installation.
\end{itemize}


\textbf{Chapter 3: Global Time}

\begin{itemize}
\item
  Notions of causal order, temporal order, and delivery order
\item
  External observers, reference clocks, and global time base
\item
  Sparse time base to view event order in a distributed real-time system
\item
  Internal clock synchronization to compensate for drift offset.
  Influence of the communication system jitter on the precision of the
  global time base.
\item
  External time synchronization, time gateways, and the Internet network
  time protocol (NTP)
\end{itemize}

\textbf{Chapter 4: Modeling Real-Time Systems}

\begin{itemize}
\item
  Introduction of a conceptual model for real-time systems
\item
  Tasks, nodes, fault-tolerant units, clusters
\item
  Simple and complex tasks
\item
  Interface placement and interface layout
\item
  Temporal control and logical control
\item
  The history state
\end{itemize}

\textbf{Chapter 5: Real-Time Entities and Images}

\begin{itemize}
\item
  Real-time entities
\item
  Observations, state and event observations
\item
  Real-time images as current picture of real-time entity, and real-time
  objects
\item
  Temporal accuracy and state estimation to improve real-time image
  accuracy
\item
  Permanence in case of race conditions and idempotency with replicated
  messages
\item
  Replica determinism to implement fault-tolerance by active redundancy
\end{itemize}

\textbf{Chapter 6: Fault Tolerance}

\begin{itemize}
\item
  Failures, Errors, and Faults
\item
  Error Detection
\item
  A Node as a Unit of Failure
\item
  Fault Tolerant Units
\item
  Reintegration of a Repaired Node
\item
  Design Diversity
\end{itemize}

\textbf{Chapter 7: Real-Time Communication}

\begin{itemize}
\item
  Real-Time Communication Requirements
\item
  Flow Control
\item
  OSI Protocols for Real-Time
\item
  Fundamental Conflicts in Protocol Design
\item
  Media-Access Protocols
\item
  Performance Comparison: ET versus TT
\item
  The Physical Layer
\end{itemize}

\textbf{Chapter 8: Time-Triggered Protocols}

\begin{itemize}
\item
  Introduction to Time-Triggered Protocols
\item
  Overview of the TTP/C Protocol Layers
\item
  The Basic CNI
\item
  Internal Operation of TTP/C
\item
  TTP/A for Field Bus Applications
\end{itemize}

\textbf{Chapter 9: Input and Output}

\begin{itemize}
\item
  The dual role of time
\item
  Agreement protocol
\item
  Sampling and polling
\item
  Interrupts
\item
  Sensors and actuators
\item
  Physical installation
\end{itemize}

\textbf{Chapter 10: Real-Time Operating Systems: OSEK and AUTOSAR}

\begin{itemize}
\item
  Task management
\item
  Interprocess communication
\item
  Time management
\item
  Error detection
\item
  OSEK and AUTOSAR
\end{itemize}

\textbf{Chapter 11: Real-Time Scheduling}

\begin{itemize}
\item
  The scheduling problem
\item
  The adversary problem
\item
  Dynamic scheduling, dynamic priority servers
\item
  Static scheduling, fixed priority servers
\end{itemize}

\textbf{Chapter 12: Validation}

\begin{itemize}
\item
  Building a Convincing Safety Case
\item
  Formal Methods
\item
  Testing
\item
  Fault Injection
\item
  Dependability Analysis
\end{itemize}

